---
title: 'Peer-graded Assignment: Course Project 1'
author: "Fabien Le Thuaut"
date: "5 October 2016"
output: html_document
---
 
#LOADING AND PREPROCESSING DATA
unzip data into a CSV
```{r unzip}
unzip("Activity monitoring data.zip")

activity <- read.csv("activity.csv")

activityday <-aggregate(steps ~ date, data=activity, sum, na.rm = TRUE)

```

##MEAN NUMBER OF STEPS TAKEN PER DAY
the data is then aggregated to calculate the mean number of steps per day
```{r agg}
activityday <-aggregate(steps ~ date, data=activity, sum, na.rm = TRUE)
```

The plot is then generated
```{r hist, echo=FALSE}
hist(activityday$steps, main = "Histogram of Steps", xlab = "Steps per Day")

```

mean value is:
```{r mean}
mean(activityday$steps)
```

media value is:
```{r median} 
median(activityday$steps)
```

##AVERAGE DAILY ACTIVITY PATTERN
Get mean data by using the aggregate function on intervals
```{r mean} 
activitystepmean <- aggregate(steps ~ interval, data=activity, mean, na.rm=TRUE)
```

and plot is to get a timeseries
```{r ts, echo=FALSE}
plot(activitystepmean$interval,activitystepmean$steps, type = "l", xlab = "interval", ylab = "steps")
```

the max number of steps is
```{r max} 
maxsteps <- max(activitystepmean$steps)
maxstepsinterval <- activitystepmean[activitystepmean$steps==maxsteps,1]
```

##IMPUTING MISSING VALUES
My strategy is to apply interval mean to NA values. I do this using the following code:
```{r Nas} 
activitynew<- activity
activityna<- is.na(activitynew$steps)
avg_interval<- tapply(activitynew$steps, activitynew$interval, mean, na.rm=TRUE, simplify = TRUE)
activitynew$steps[nas] <- avg_interval[as.character(activitynew$interval[activityna])]

activitynewday <- aggregate(steps ~ date, data=activitynew, sum, na.rm = TRUE)
```

I am then able to output the following graph. Similar to the first histogram but without the NA values
```{r hist2, echo=FALSE}
hist(activitynewday$steps, main = "Histogram of Steps", xlab = "Steps per Day")
```

I then recalculate the mean and median values:
```{r meanmedian} 
mean(activitynewday$steps)
median(activitynewday$steps)
```

##DIFFERENCES IN ACTIVITY PATTERNS BETWEEN WEEKDAY AND WEEKENDS
```{r weekday/end} 
I then classify each row as Weekday or Weekend using the date column with the following code:
activity3 <- cbind(activitynew,weekdays(as.Date(activitynew$date)))
names(activity3)[4] <- "DayOfWeek"

weekdays <- subset(activity3,as.character(activity3[DayOfWeek]) %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"))
weekends <- subset(activity3,activity3$DayOfWeek %in% c("Saturday","Sunday"))

weekdayint<-aggregate(steps ~ interval, data=weekdays, mean, na.rm=TRUE)
weekendint<-aggregate(steps ~ interval, data=weekends, mean, na.rm=TRUE)
```

and lastly output a pannel plot:
```{r hist2, echo=FALSE}
par(mfrow=c(2,1))
plot(weekdayint$interval,weekdayint$steps, type = "l", xlab = "interval", ylab = "steps",main = "Weekdays")
plot(weekendint$interval,weekendint$steps, type = "l", xlab = "interval", ylab = "steps",main = "Weekends")
```